---
title: "Batch Processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Batch Processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(LeafInjuryR)
out <- file.path(tempdir(), "leaf_batch")
b <- analyze_leaf_batch(leaf_example_images(), output_dir = out,
                        crop = "auto", progress = FALSE)
b
```

`images` may also be a directory (`analyze_leaf_batch("photos/")`). All
images use the same settings, which are stored once in `b$parameters`.

## Outputs

```{r}
list.files(out, recursive = TRUE)
head(read.csv(file.path(out, "results.csv"))[, c("file", "healthy_percent",
                                                   "injured_percent", "quality_flag",
                                                   "status", "elapsed_sec")])
```

* `results.csv`: one row per image (CSV is the default format for
  interoperability and reproducibility).
* `parameters.csv` and `analysis_report.txt`: methods and parameters.
* `masks/`: leaf mask, injury mask and colour class map (PNG).
* `overlays/`: translucent overlays for visual auditing.

## Robustness

A failing image never stops the batch: it gets `status = "failed"` and an
`error_message`. Images are not kept in memory. Always inspect rows with
`quality_flag = "warning"`: warnings do not change results.

```{r}
bad <- tempfile(fileext = ".jpg"); writeLines("not an image", bad)
b2 <- analyze_leaf_batch(c(leaf_example_images()[1], bad), progress = FALSE)
b2$results[, c("file", "status", "error_message")]
```

## Performance

```{r, eval = FALSE}
benchmark_leaf_analysis(leaf_example_images(), times = 3)
```

Time grows roughly linearly with pixel count (about 1-2 s per 0.75-megapixel
image on a typical computer). Images of 1-4 megapixels with the leaf filling
20-70 % of the frame are usually sufficient; `crop = "auto"` reduces the
analysed area. Processing is sequential in version 0.1.0.
