---
title: "Batch Processing and Reports"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Batch Processing and Reports}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(LeafInjuryR)
imgs <- system.file("extdata", c("leaf_example_01.jpg", "leaf_example_02.jpg"),
                    package = "LeafInjuryR")
out <- file.path(tempdir(), "leaf_batch")
b <- analyze_leaf_batch(imgs, output_dir = out, crop = "auto", progress = FALSE)
b
b$summary
list.files(out, recursive = TRUE)
```

`images` may also be a folder: `analyze_leaf_batch("photos/", output_dir = "results")`.

## Reports (generated automatically)

All three come from the same results table:

* `report.html` - self-contained interactive report (opens offline, no R
  session needed): executive summary, statistics, comparative chart,
  sortable table, gallery of original and processed images, parameters and
  MD5 checksums of every input file.
* `results.xlsx` - sheets *Resultados*, *Resumo* and *Parametros*.
* `results.csv` - one row per image, one column per variable.

## Robustness

A failing image never stops the batch:

```{r}
bad <- tempfile(fileext = ".jpg"); writeLines("not an image", bad)
b2 <- analyze_leaf_batch(c(imgs[1], bad), progress = FALSE)
b2$results[, c("file", "status", "error_message")]
```

Rows with `quality_flag = "warning"` should be inspected; warnings never
change results. Batch analysis uses automatic segmentation; images that need
manual correction can be reopened individually.
