---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5.5)
```

LeafInjuryR quantifies the percentage of leaf area showing visual injury
patterns in photographs of detached leaves on a contrasting background, with
classical, auditable image processing that runs entirely on your computer.

## Installation

```{r, eval = FALSE}
install.packages("BiocManager")
BiocManager::install("wilhanvalasco/LeafInjuryR")   # installs EBImage too
install.packages(c("shiny", "bslib", "DT"))          # graphical interface
```

## The six functions

| Function | Purpose |
|---|---|
| `analyze_leaf()` | single-image analysis |
| `analyze_leaf_batch()` | batch analysis with HTML, Excel and CSV reports |
| `crop_leaf()` | automatic and manual cropping |
| `segment_leaf()` | automatic and manually corrected segmentation |
| `validate_leaf()` | quality control and validation against reference masks |
| `run_leafinjury_app()` | local Shiny interface |

## Analyse one image

```{r}
library(LeafInjuryR)
f <- system.file("extdata", "leaf_example_02.jpg", package = "LeafInjuryR")
res <- analyze_leaf(f, crop = "auto", multiclass = TRUE)
res
plot(res)
```

`injured_percent` = injured pixels / leaf-mask pixels x 100. Areas are in
pixels; cm^2 are reported only when a scale is given (`pixels_per_cm`).

## Quality control

```{r}
validate_leaf(res)
```

Without manual reference masks only quality indicators are shown; accuracy
metrics require ground truth (see `vignette("method-validation")`).

## Graphical interface

```{r, eval = FALSE}
run_leafinjury_app()
```

The app has four areas - Imagem, Segmentação, Análise and Resultados - and
calls the same functions.
