---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5.5)
```

LeafInjuryR quantifies the percentage of leaf area showing visual injury
patterns in photographs of detached leaves on a contrasting background. It
uses classical, auditable image processing and runs entirely on your computer.

## Installation

EBImage is distributed by Bioconductor, so install it first:

```{r, eval = FALSE}
install.packages("BiocManager")
BiocManager::install("EBImage")
# then, from the package source:
install.packages("LeafInjuryR_0.1.0.tar.gz", repos = NULL, type = "source")
# optional, for the graphical interface:
install.packages(c("shiny", "bslib", "DT"))
```

## Analyse one image

```{r}
library(LeafInjuryR)
files <- leaf_example_images()
res <- analyze_leaf(files[2], crop = "auto")
res
```

The four audit panels show the analysed region, the leaf mask, the tissue
classification and the translucent overlay:

```{r}
plot(res)
```

The tabular result is a one-row data frame:

```{r}
res$metrics[, c("file", "leaf_pixels", "healthy_pixels", "injured_pixels",
                "healthy_percent", "injured_percent", "quality_flag")]
```

`injured_percent` is always computed relative to **leaf pixels**, never to the
whole image.

## Multiclass classification

```{r}
res_mc <- analyze_leaf(files[2], crop = "auto", multiclass = TRUE)
res_mc$metrics[, c("healthy_percent", "chlorotic_percent",
                   "necrotic_percent", "other_percent", "injured_percent")]
```

By construction, `injured = chlorotic + necrotic + other`.

## Quality control and parameters

```{r}
res$quality$quality_flag
res$quality$quality_messages
analysis_report(res)
```

Every result stores the parameters needed to reproduce it
(`get_analysis_parameters(res)`), including package version, thresholds
actually used, morphology, quality-control limits and the random seed.

## Graphical interface

```{r, eval = FALSE}
run_leafinjury_app()
```

## Important

The example images are functional examples, **not** ground truth. Working
software is not a validated method: see `vignette("method-validation")`.
