---
title: "Cropping, Segmentation and Manual Correction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cropping, Segmentation and Manual Correction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
```

```{r}
library(LeafInjuryR)
f <- system.file("extdata", "leaf_example_02.jpg", package = "LeafInjuryR")
```

## Cropping

`crop_leaf()` never resamples: pixels keep their original resolution.

```{r}
img <- crop_leaf(f, mode = "auto")
attr(img, "crop_info")$bbox
img_m <- crop_leaf(f, mode = "manual", x = c(200, 900), y = c(60, 650))
dim(img_m)
```

## Two separate problems

1. **Leaf segmentation** - which pixels are leaf, healthy or not. Necrotic
   margins must be included, otherwise injury is underestimated.
2. **Tissue classification** - which *leaf* pixels look healthy or injured,
   computed only inside the leaf mask.

```{r}
seg <- segment_leaf(img, multiclass = TRUE)
seg
plot(seg)
seg$details$segmentation$strategy
```

The default leaf segmentation combines CIELAB chroma (green and yellow
tissue) with darkness (dark necrosis). Excess-green indices exclude brown
necrosis and are offered only for comparison.

The default tissue rule (`"lab"`) uses fixed thresholds, identical for all
images, which is preferable when comparing treatments. `"relative"` measures
the distance to the leaf's own dominant colour and is intended for naturally
pigmented or senescent leaves, where the absence of green is not a valid
injury criterion.

## Manual correction

Corrections are rows of a data frame (brush strokes or rectangles). They are
applied in order, stored in the result and therefore reproducible.

```{r}
fix <- data.frame(action = c("healthy", "remove_leaf"),
                  shape = c("rect", "circle"),
                  xmin = c(40, NA), xmax = c(140, NA), ymin = c(220, NA), ymax = c(300, NA),
                  x = c(NA, 20), y = c(NA, 20), radius = c(NA, 12))
seg2 <- segment_leaf(img, mode = "manual", corrections = fix, segmentation = seg)
c(before = seg$metrics$injured_percent, after = seg2$metrics$injured_percent)
seg2$corrections
```

In the Shiny app the same corrections are made with a brush (click) or a
rectangle, and the injury percentage updates immediately.

## Scope

LeafInjuryR quantifies visual colour patterns. It does not diagnose
pathogens, diseases, nutritional disorders or phytotoxicity.
