---
title: "Image Segmentation and Tissue Classification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Image Segmentation and Tissue Classification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

```{r}
library(LeafInjuryR)
img <- read_leaf(leaf_example_images()[2])
```

## Two different problems

1. **Leaf segmentation** (`segment_leaf()`): which pixels are leaf,
   whatever their health status. Necrotic and chlorotic tissue *must* be
   included, otherwise injury is underestimated.
2. **Injury classification** (`classify_leaf_tissue()`): which *leaf*
   pixels look healthy or injured. It runs only inside the leaf mask.

A visually acceptable mask is not evidence of accuracy; see
`vignette("method-validation")`.

## Segmentation methods

```{r}
cmp <- compare_segmentation_methods(img, methods = c("hsv", "lab", "exg", "otsu", "auto"))
cmp
plot(cmp)
```

Note how `exg` (excess green) excludes the brown necrotic margin: indices
built to detect *green vegetation* are not appropriate to delimit an injured
leaf. The `auto` method combines CIELAB chroma (captures green and yellow
tissue) with darkness (captures dark necrosis), and infers which side of each
threshold is background from the image border. Agreement between methods is
not accuracy; no method is declared best without ground truth.

The decision taken by `auto` is recorded:

```{r}
seg <- segment_leaf(img, "auto")
seg$details[c("strategy", "accepted_features", "candidate_thresholds")]
```

## Morphology and connected components

```{r}
seg$components
seg$holes[c("policy", "hole_pixels", "filled_pixels", "perforation_candidate_pixels")]
```

Components removed as noise or as external objects are listed with the
reason. Holes are filled only when their colour differs from the
background; background-coloured holes are kept outside the leaf and reported
as *perforation candidates* (experimental) - they are never counted as injury.

## Colour features

No single channel is assumed to be universally superior. The features of
leaf pixels can be explored directly:

```{r}
feat <- leaf_color_features(img, seg$mask, max_pixels = 20000)
round(sapply(feat[, c("L", "a", "b", "chroma", "lab_hue", "hsv_hue",
                      "saturation", "exg", "exr", "exgr")], quantile,
             c(0.05, 0.5, 0.95)), 2)
```

## Tissue methods

```{r}
compare_tissue_methods(img, seg$mask)$summary[, c("tissue_method", "injured_percent")]
```

The spread between rules on the same leaf mask shows that the **tissue
thresholds are the main source of uncertainty** in the injury estimate. The
default (`"lab"`, fixed thresholds) applies identical rules to every image,
which is preferable for comparing treatments; thresholds must be calibrated
for each imaging setup against reference masks.

## Normalisation

```{r}
img_n <- normalize_leaf_image(img)
attr(img_n, "normalization")$white_balance_gains
```

Corrections are estimated from background pixels only and clipped to a
conservative range. `analyze_leaf(normalize = TRUE)` keeps both
`result$original` and `result$normalized` for auditing.

## Scope

Colour changes can have many biological causes. LeafInjuryR quantifies visual
patterns; it does not diagnose pathogens, diseases, nutritional disorders or
phytotoxicity.
