---
title: "Method Validation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Method Validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
```

## Software that works is not a validated method

Passing unit tests shows computational correctness (the denominator is leaf
pixels, masks are consistent, percentages add up). It does **not** show that
injury percentages agree with expert assessment for a given species, camera
and lighting. That requires **ground truth**: masks drawn manually.

The example images shipped with the package have no reference masks and must
not be used as ground truth.

## Reference masks

For each validation photograph draw, in GIMP, Krita, ImageJ/Fiji, CVAT or
similar:

* a **leaf mask** (white = leaf, black = everything else);
* an **injury mask** (white = injured tissue).

Use the size of the original photograph, save as PNG or TIFF, document the
criteria used and, ideally, use more than one annotator.

## Metrics

| Metric | Formula |
|---|---|
| IoU (Jaccard) | TP / (TP + FP + FN) |
| Dice (= F1 for masks) | 2TP / (2TP + FP + FN) |
| Sensitivity | TP / (TP + FN) |
| Specificity | TN / (TN + FP) |
| Precision | TP / (TP + FP) |

`validate_leaf()` evaluates separately (A) leaf x background and (B)
healthy x injured inside the leaf, and reports the error of the injury
percentage. Without references it reports quality indicators only.

## Worked example (synthetic image with known truth)

```{r}
library(LeafInjuryR)
syn <- LeafInjuryR:::make_synthetic_leaf(chlorotic = TRUE)
res <- analyze_leaf(syn$image)
v <- validate_leaf(res, reference_leaf_mask = syn$leaf_mask,
                   reference_injury_mask = syn$injury_mask)
v
plot(v)
```

Synthetic images only test the software.

## Recommended protocol

1. Collect a representative validation set (species, cultivars, injury from
   0 to high, cameras, days).
2. Tune thresholds (`params`) on a calibration subset; report performance on
   an independent test subset.
3. Report IoU/Dice for (A) and (B), sensitivity, specificity and the error
   of injury percent (e.g. Bland-Altman), with number of images and
   annotators.
4. Freeze the configuration (`result$parameters$config`) and package version.
