---
title: "Method Validation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Method Validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Software that works is not a validated method

LeafInjuryR runs, produces masks and numbers, and passes its unit tests.
This demonstrates *computational* correctness (e.g. the denominator is leaf
pixels, masks are consistent, percentages add up). It does **not**
demonstrate that the injury percentages agree with expert assessment for a
given species, camera, background and lighting. That requires comparison
with **ground truth**: masks drawn manually by trained people.

The two example images shipped with the package have no reference masks and
must not be used as ground truth.

## Ground truth

For each validation image produce, in an image editor (GIMP, Krita,
ImageJ/Fiji, CVAT, Label Studio, ...):

* `reference_leaf_mask`: white = leaf, black = everything else;
* `reference_injury_mask`: white = injured tissue, black = everything else.

Rules: same width and height as the original photograph (masks are never
resized silently); save as PNG or TIFF (JPEG compression changes edges);
draw on the original, not on a cropped or resized copy; document the
criteria used to call tissue injured; ideally have more than one annotator
to estimate inter-annotator variability.

Then list them in a CSV (template:
`system.file("extdata", "ground_truth_template.csv", package = "LeafInjuryR")`):

```
image,reference_leaf_mask,reference_injury_mask
images/leaf01.jpg,masks/leaf01_leaf.png,masks/leaf01_injury.png
```

and read it with `read_ground_truth()`.

## Metrics

With TP, FP, FN and TN pixel counts (positive = leaf, or = injured):

| Metric | Formula | Meaning |
|---|---|---|
| IoU (Jaccard) | TP / (TP + FP + FN) | overlap relative to union |
| Dice | 2TP / (2TP + FP + FN) | overlap; equals F1 for masks |
| Sensitivity (recall) | TP / (TP + FN) | share of true positives found |
| Specificity | TN / (TN + FP) | share of true negatives kept |
| Precision | TP / (TP + FP) | share of detections that are correct |
| Accuracy | (TP + TN) / N | inflated when negatives dominate |

Evaluate separately:

* **A) leaf x background** - is the leaf delimited correctly?
* **B) healthy x injured** - within the leaf, is tissue classified correctly?
  `evaluate_leaf_analysis()` restricts (B) to the leaf so that background
  pixels do not inflate specificity or accuracy.

Also report the error of the final number, `predicted - reference` injured
percent, and its agreement across images (e.g. Bland-Altman analysis,
concordance correlation).

## Worked example (synthetic, known truth)

```{r}
library(LeafInjuryR)
dir <- file.path(tempdir(), "gt_demo"); dir.create(dir, showWarnings = FALSE)
syn <- make_synthetic_leaf(chlorotic = TRUE)
write_image_png(syn$image, file.path(dir, "leaf.png"))
write_mask_png(syn$leaf_mask, file.path(dir, "leaf_leafmask.png"))
write_mask_png(syn$injury_mask, file.path(dir, "leaf_injurymask.png"))

man <- ground_truth_manifest(file.path(dir, "leaf.png"),
                             file.path(dir, "leaf_leafmask.png"),
                             file.path(dir, "leaf_injurymask.png"))
validate_ground_truth(man)[, c("image", "label", "iou", "dice", "sensitivity",
                               "specificity", "precision", "injured_percent_error")]
```

Synthetic images only check the software; they say nothing about real
photographs.

## Recommended protocol

1. Collect a validation set representative of the conditions of use
   (species, cultivars, injury levels from 0 to high, cameras, days).
2. Split into a *calibration* subset (to tune thresholds in `leaf_config()`)
   and an independent *test* subset (to report performance). Never report
   performance on the images used for tuning.
3. Report IoU/Dice for (A) and (B), sensitivity, specificity, and the
   error of injured percent, with the number of images and annotators.
4. Freeze the configuration (`get_analysis_parameters()`) and package
   version used in the study.
